\section{Heyting-valued Sets}

\def\h{\mathcal{H}}

\subsection{Preliminaries}

\begin{definition}
  A \emph{Heyting-valued set \(A\)} is a carrier set \(\carr A\) equipped with a  ``Heyting-valued relation'' \(\eq[A]{-}{-} : A×A → \h\) satisfying
  \begin{gather}
    \label{hvs:symm}  \eq x y ≤ \eq y x\\
    \label{hvs:trans} \eq x y ∧ \eq y z ≤ \eq x z
  \end{gather}

  Define \(\e x ≔ \eq x x\).
\end{definition}

\begin{definition}
  A \emph{Heyting-valued morphism \(f : A ↬ B\)} is a ``Heyting-valued functional relation'', meaning it is a map \(\carr f : \carr A × \carr B → \h\) satisfying
  \begin{gather}
    \label{hvm:str}    f(x,y) ≤ \e x ∧ \e y\\
    \label{hvm:well}   \eq {x'} x ∧ f(x, y) ∧ \eq y {y'} ≤ f(x', y')\\
    \label{hvm:unique} f(x, y) ∧ f(x, y') ≤ \eq y {y'}\\
    \label{hvm:total}  \e x ≤ ⋁_{y ∈ B}f(x, y)
  \end{gather}
\end{definition}

\begin{fact}
  The following hold.

  \begin{enumerate}
  \item \(f(x,y)∧\eq x {x'} = f(x',y)∧\eq x {x'}\)
  \item \(\e x = ⋁_{y∈A}\eq x y\)
  \end{enumerate}
\end{fact}

\begin{construction}
  The product \(A×B\) is given by \(\carr A × \carr B\) with equality
  \[ \eq{\p{x,y}}{\p{x',y'}} ≔ \eq x {x'} ∧ \eq y {y'}\text. \]

  The projection \(π : A×B ↬ A\) is defined by
  \[ \p{\p{x,y}, x'} ↦ \eq x {x'} ∧ \e y\text. \]
  The other projection is defined similarly
\end{construction}
\begin{proof}
  Symmetry and transitivity of the equality follow clearly.

  Likewise, the projection is clearly strict and well-defined.
  Uniqueness holds via
  \begin{align*}
    π(\p{x,y},x') ∧ π(\p{x,y},x'')
    &= \eq x {x'} ∧ \e y ∧ \eq x {x''} ∧ \e y\\
    &= \eq x {x'} ∧ \eq x {x''} ∧ \e y\\
    &≤ \eq {x'} {x''} ∧ \e y\\
    &≤ \eq {x'} {x''}\text.
  \end{align*}
  Totality holds via
  \begin{align*}
    \e{\p{x,y}} = \e x ∧ \e y = ⋁_{x'∈A}\eq x {x'} ∧ \e y\text.
  \end{align*}
\end{proof}

In the following, I will not write out these proofs. A standard reference is
TODO BorceuxV3.

\begin{construction}
  The \emph{discrete Heyting-valued set \(ΔA\) on \(A\)} has \(A\) as it's carrier and equality defined by \(\p{x,y} ↦ ⋁\set{⊤}{x=y}\).
\end{construction}

\begin{fact}
  The empty type \(∅\) is given by the empty carrier.

  The unit type \(\mb 1\) is given by \(\{\ast\}\) with \(\e\ast = ⊤\). This is the discrete Heyting-valued set on \(\{\ast\}\).

  The natural numbers object is the discrete Heyting-valued set on \(ℕ\).

  The rational numbers object is the discrete Heyting-valued set on \(ℚ\).
\end{fact}
\begin{note}
  The real numbers object is not necessarily discrete.
\end{note}

\begin{construction}
  The \emph{sum \(A+B\)} is given by \(\carr A+\carr B\) with equality
  \begin{align*}
    &\eq {\inl x} {\inl x'} ≔ \eq[A] x {x'}\\
    &\eq {\inl x} {\inr y'} ≔ ⊥\\
    &\eq {\inr y} {\inr x'} ≔ ⊥\\
    &\eq {\inr y} {\inr y'} ≔ \eq[B] y {y'}
  \end{align*}

  The inclusion \(ι : A ↬ A+B\) is given by \(\p{x,t} ↦ \eq {\inl x} t\).
  The other inclusion is defined similarly.
\end{construction}

\begin{construction}
  The \emph{subobject classifier \(Ω\)} is given by \(\h\) with \(∧\) as equality.
\end{construction}

\begin{construction}
  The \emph{exponential object \(B^A\)} is given by
  \[ \set{f : \carr A × \carr B → \h}{\ref{hvm:str}-\ref{hvm:unique}} \]
  with equality
  \[ \p{f,g} ↦ ⋀_{x∈A}⋁_{y∈B}f(x,y) ∧ g(x,y)\text. \]
\end{construction}
\begin{internal}
  The equality in the internal language is \(\for{x∈A}{\exist{y∈B}{f(x, y) ∧ g(x, y)}}\) which can be rephrased as \(\for{x∈A}{f(a) = g(a)}\).
  Thus equality models function extensionality, which is know to hold in Grothendieck toposes.
\end{internal}

\begin{fact}
  The \emph{power-object \(𝒫A = Ω^A\)} can be given as the discrete Heyting-valued set on \(\set{P : \carr A → Ω}{\eq {x'} x ∧ P(x) ≤ P(x')}\).
\end{fact}

\begin{construction}
  The \emph{restriction \(A\res p\)} is given by \(\set{x ∈ \carr A}{\e x \between p}\) with equality \(\eq[A\res p] a {a'} ≔ \eq[A] a {a'} ∧ p\).
\end{construction}
\begin{note}
  Taking the above subset of the carrier has the effect of removing ``junk'' elements, i.~e. elements at stage \(⊥\).
  This may or may not be necessary and may or may not improve the practical application of this construction. TODO: figure out if it does
\end{note}

\begin{construction}
  Given a Heyting-valued set \(A\) and a relation \(R : \carr A × \carr A → \h\) satisfying
  \begin{gather}
    \e x ≤ R x x\\
    R x y ≤ R y x\\
    R x y ∧ R y z ≤ R x z
  \end{gather}
  we can form the \emph{quotient \(\quot A R\)} with carrier \(\carr A\) and equality \(R\).
\end{construction}
\begin{note}
  Subquotients can be formed similarly, by merely dropping the first condition on the relation.
  If instead \(R x x ≤ \e x\) holds the above construction constructs the subset given by the predicate \(R x x\).
\end{note}

\subsection{CwF structure}

To define what is a family we first define a ``universe''.
%Call a class equipped with a Heyting-valued relation satisfying \ref{hvs:symm} and \ref{hvs:trans} a \emph{Heyting-valued class}. The above constructions generalise trivially to Heyting-valued classes.

\begin{construction}
  Up to size issues, the class of all Heyting-valued sets \(\hvsets\) with equality \(\p{A,B} ↦ ⋁\set{p∈\h}{A\res p = B\res p}\) is a Heyting-valued set.

  We will call it the \emph{universe \(𝒰\)}.
\end{construction}
\begin{note}
  The equality above is a strict equality in the meta-theory, in which two Heyting-valued sets are equal when their carriers are equal, and their equality relations are equal as functions.
\end{note}

\begin{construction}
  Given a  family \(\p{Bₐ}_{a∈A}\) of Heyting-valued sets, such that \(Bₐ\res{\eq a {a'}} = B_{a'}\res{\eq a {a'}}\), the \emph{gluing \(⋁B\)} has carrier
  \[ \set{b ∈ \carr{Bₐ\res{\e a}}}{a ∈ \carr A} = ⋃_{a ∈ \carr A}\carr{Bₐ\res{\e a}} \]
  and equality
  %\[ \p{b, b'} ↦ ⋁_{a,a' ∈ A} b ∈ Bₐ\res {\eq a {a'}} ∧ b' ∈ B_{a'}\res{\eq a {a'}}
  %  ∧ \eq a {a'} ∧ \eq[Bₐ\res {\eq a {a'}}] b {b'}\text. \]
  %\[ \p{b, b'} ↦ ⋁_{a ∈ A} b,b' ∈ Bₐ\res {\e a} ∧ \eq[Bₐ\res {\e a}] b {b'}\text. \]
  \[
    \eq[⋁B] b {b'} ≔
    ⋁\set{\eq[Bₐ\res {\e a}] b {b'}}
    {a ∈ \carr A\text{, } b,b' ∈ \carr{Bₐ\res {\e a}}} \text.
  \]
\end{construction}
\begin{internal}
  Elements of the gluing should obviously come from the elements of individual sets \(Bₐ\), ``modulo equality''. Since we are working with Heyting-valued sets we know we can move the quotient from the carrier set to the equality relation, as we do here.

  Internally, the equality relation would look something like
  \[ \exist{a ∈ A}{b,b' ∈ Bₐ ∧ \eq[Bₐ] b {b'}} \]
  and this should be our working intuition for the equality.

  Of note is that we omit elements of \(Bₐ\) that don't exist anywhere on \(\e a\). This does not really matter, as those elements would not exist in the gluing, but this gives a simpler presentation and might be useful later.
\end{internal}
\begin{proof}
  Let \(b₀, b₁, b₂ ∈ ⋁B\). Then by construction \(bᵢ ∈ B_{aᵢ}\res{\e {aᵢ}}\) for some \(aᵢ ∈ A\).
  \begin{align*}
    \eq {b₀} {b₁}
    &= ⋁_{a ∈ A} b₀,b₁ ∈ Bₐ\res {\e a} ∧ \e a ∧ \eq[Bₐ\res {\e a}] {b₀} {b₁}\\
    &≤ ⋁_{a ∈ A} b₁,b₀ ∈ Bₐ\res {\e a} ∧ \e a ∧ \eq[Bₐ\res {\e a}] {b₁} {b₀}\\
    &= \eq {b₁} {b₀}
  \end{align*}
  \begin{align*}
    \eq {b₀} {b₁} ∧ \eq {b₁} {b₂}
    &= ⋁_{a,a' ∈ A} b₀,b₁ ∈ Bₐ\res {\e a} ∧ \e a ∧ \eq[Bₐ\res {\e a}] {b₀} {b₁}
    ∧ b₁,b₂ ∈ B_{a'}\res {\e {a'}} ∧ \e {a'} ∧ \eq[B_{a'}\res {\e {a'}}] {b₁} {b₂}\\
    &= ⋁_{a ∈ A} b₀,b₁,b₂ ∈ Bₐ\res {\e a} ∧ \e a
    ∧ \eq[Bₐ\res {\e a}] {b₀} {b₁} ∧ \eq[Bₐ\res {\e a}] {b₁} {b₂}\\
    &≤ ⋁_{a ∈ A} b₀,b₂ ∈ Bₐ\res {\e a} ∧ \e a
    ∧ \eq[Bₐ\res {\e a}] {b₀} {b₂}\\
    &= \eq {b₀} {b₂}\qedhere
  \end{align*}
\end{proof}

\begin{definition}
  A \emph{family \(B\) over \(A\)} is a map \(B : A → 𝒰\) satisfying
  \begin{align*}
    &φ(x)\res{\e x} = φ(x)\text{ and}\\
    &φ(x)\res{\eq x {x'}} = φ(x')\res{\eq x {x'}}\text.
  \end{align*}

  We denote the collection of families over \(A\) by \(\fam A\).
  The first property is called \emph{canonicity} and the second is called \emph{coherence}.
\end{definition}
TODO: is canonicity necessary?

\begin{construction}
  The isomorphism between morphisms \(A ↬ 𝒰\) and families over \(A\) is given by
  \mordef{\p{-}^\flat}{\p{A ↬ 𝒰}}{\fam A}{B}{\p{a ↦ ⋁\p{X\res{B(a,X)}}_{\mkern-6muX ∈ 𝒰}}}
  and
  \mordef{\tilde{\p{-}}}{\fam A}{\p{A ↬ 𝒰}}{B}{\p{\p{a, X} ↦ \e a ∧ \eq[𝒰] {B(a)} X}\text.}
\end{construction}
\begin{proof}
  The relation \(\tilde B\) is obviously a morphism \(A ↬ 𝒰\) for \(B ∈ \fam A\).

  For the other direction fix \(a ∈ A\) and \(B : A ↬ 𝒰\).

  Because \(B\) is a morphism we have \(B(a,X\res{B(a,X)}) = B(a,Y\res{B(a,Y)})\) iff \({X\res{B(a,X)}} = {Y\res{B(a,Y)}}\).

  This means we can rewrite the gluing as being over the set \(\set{B(a, X)}{X ∈ 𝒰}\).
  This means that the carrier is a small union of small sets, and is thus small.
  Therefore it's a Heyting-valued set and thus an element of \(𝒰\), so this indeed defines a map \(A → 𝒰\).

  Take now another \(a'\) to show the family coherence property.
  The restriction \(B^\flat(a)\res{\eq a {a'}}\) has as carrier
  \[ \set{x ∈ X}{X ∈ 𝒰 ∧ \p{\e x ∧ B(a, X)} \between \eq a {a'}}\text. \]
  The intersecting condition is equivalent to
  \[ ⊥ < \e x ∧ f(a,X) ∧ \eq a {a'}\text, \]
  which is obviously equivalent to \(\p{\e x ∧ f(a', X)} \between \eq a {a'}\).
  Thus we have \({B^\flat(a)\res{\eq a {a'}} = B^\flat(a')\res{\eq a {a'}}}\)
  and \(B^\flat\) is a family over \(A\).

  It remains to show the maps are inverses.
  First let \(a ∈ A\) and \(B : \fam A\).
  \begin{align*}
    {\tilde{B}}^\flat(a)
    &= ⋁_{X ∈ 𝒰}\p{X\res{\tilde B(a,X)}}\\
    &= ⋁_{X ∈ 𝒰}\p{X\res{\e a ∧ \eq {B(a)} X}}\\
    &= ⋁_{X ∈ 𝒰}\p{\p{X\res{\eq {B(a)} X}}\res{\e a}}\\
    &= ⋁_{X ∈ 𝒰}\p{\p{B(a)\res{\eq {B(a)} X}}\res{\e a}}\\
    &= B(a)\res{\e a}\\
    &= B(a)
  \end{align*}

  Now let \(a ∈ A\), \(X ∈ 𝒰\), and \(B : A ↬ 𝒰\).
  \begin{align*}
    \tilde{B^\flat}(a, X)
    &= \e a ∧ \eq {B^\flat(a)} X\\
    &= \e a ∧ ⋁_{Y ∈ 𝒰}\eq {Y\res{B(a,Y)}} X\\
    &= \e a ∧ ⋁_{Y ∈ 𝒰}\eq Y X ∧ B(a, X)\\
    &= B(a, X)\qedhere
  \end{align*}
\end{proof}
% \begin{proof}
%   Take \(a ∈ A\). The type \(φ(a)\) is the gluing of the family
%   \(\p{X\res{f(a,X)}}_{\mkern-6muX ∈ 𝒰}\), that is the set
%   \(\set{x ∈ X}{X ∈ 𝒰 ∧ \e x \between f(a, X)}\) with equality
%   \[ \p{x,y} ↦ ⋁_{X∋x,\\Y∋y}f(a, X) ∧ f(a, Y) ∧ \eq[X\res{\eq X Y}] x y\text. \]

%   This gluing is small since the collection \(I ≔ \{X\res{f(a,X)} ∈ \hvsets\}\) is a
%   set.
%   If \(f(a,X\res{f(a,X)}) = f(a,Y\res{f(a,Y)})\) then
%   \(\eq[𝒰] {X\res{f(a,X)}} {Y\res{f(a,Y)}}\).
%   Therefore the collection \(I\) has at most \(\h\) many elements.

%   Take another \(a' ∈ A\). The restriction \(φ(a)\res{\eq a {a'}}\) has as
%   carrier
%   \[ \set{x ∈ X}{X ∈ 𝒰 ∧ \e x ∧ f(a,x) \between \eq a {a'}}\text. \]
%   The condition \(\e x ∧ f(a, X) \between \eq a {a'}\) is equivalent to
%   \[ ⊥ < \e x ∧ f(a, X) ∧ \eq a {a'} \text, \]
%   so it is also equivalent to \(\e x ∧ f(a', X) \between \eq a {a'}\).

%   Thus, \(φ(a)\res{\eq a {a'}} = φ(a')\res{\eq a {a'}}\).

%   In the other direction taking a map \(φ : A → 𝒰\) we define
%   \(f(a, X) ≔ \e a ∧ \eq X {φ(a)}\). This is obviously a morphism
%   \(f : A ↬ 𝒰\).

%   If \(f = g : A ↬ 𝒰\), then for \(a ∈ A\) we have
%   \[ φ(a) = ⋁\p{X\res{f(a, X)}} = ⋁\p{X\res{g(a, X)}} = γ(a)\text. \]

%   Let \(a ∈ A\), \(φ : A → 𝒰\), and \(f\) the corresponding morphism. Let \(γ\)
%   denote the corresponding map to \(f\).
%   \begin{align*}
%     γ(a)
%     &≔ ⋁\p{X\res{f(a, X)}}\\
%     &= ⋁\p{X\res{\e a ∧ \eq X {φ(a)}}}\\
%     &= ⋁\p{X\res{\eq X {φ(a)}}}\res{\e a}\\
%     &= ⋁\p{φ(a)}\res{\e a}\\
%     &= φ(a)
%   \end{align*}

%   Let now \(f : A ↬ 𝒰\), \(φ\) the corresponding map to \(f\), and \(g\) be the
%   corresponding morphism to \(γ\). It remains to show that \(g = f\). For
%   \(a ∈ A\) and \(X ∈ 𝒰\) we have
%   \begin{align*}
%     g(a, X)
%     &= \e a ∧ \eq X {φ(a)}\\
%     &= \e a ∧ ⋁_{Y ∈ 𝒰} \eq X {Y\res{f(a, Y)}}\\
%     &= \e a ∧ ⋁_{Y ∈ 𝒰} \eq X Y ∧ f(a, Y)\\
%     &= \e a ∧ f(a, X)\\
%     &= f(a, X)
%   \end{align*}
% \end{proof}


\begin{construction}
  Up to size issues the category of Heyting-valued sets is a category with families.

  Let \(Γ\) and \(Δ\) be Heyting-valued sets, \(A\) a family over \(Γ\), and \(γ : Δ ↬ Γ\).
  \begin{itemize}
  \item \(\ctx ≔ \hvsets\)
  \item \(\ty Γ ≔ \fam Γ\)
  \item \(\ty γ ≔ A ↦ \p{\tilde A ∘ γ}^♭\)
  \item \(\tm Γ A ≔ \set
    {\p{\sigma{g ∈ \carr Γ}{\carr {A(g)}}} → \h}
    {\ref{hvm:str}-\ref{hvm:total}}\)
  \item \(\tm γ A ≔ f ↦ f ∘ γ\)
  %\item \(\ext Γ A ≔ \p{\sigma{g ∈ \carr Γ}{\carr{A(g)}},
  %    \p{\p{g, a},\p{g',a'}} ↦ \eq[Γ] g {g'} ∧ \eq[A(g)] a {a'}}\)
  \item \(\carr{\ext Γ A} ≔ \sigma{g ∈ \carr Γ}{\carr{A(g)}}\)
  \item \(\eq[\ext Γ A] {\p{g, a}} {\p{g',a'}} ≔ \eq[Γ] g {g'} ∧ \eq[A(g)] a {a'}\)
    NOTE: gluing?
  \item \(\ext γ t ≔ \p{d,\p{g,a}} ↦ γ(d, g) ∧ t(d, a)\)
  \end{itemize}
  Call this category with families \(\hvcwf\)
\end{construction}

\begin{construction}
  \(\hvcwf\) has dependent sums. Let \(Γ : \ctx\), \(A : \ty Γ\), and \(B : \ty \ext Γ A\).
  \begin{itemize}
  %\item \(Σ Α Β (g) ≔ \p{\sigma{a ∈ \carr{A(g)}}{\carr{B(g, a)}},
  %    \p{\p{a,b},\p{a',b'}} ↦ \eq[A(g)] a {a'} ∧ \eq[B(g,a)] b {b'}}\)
  \item \(\carr{Σ Α Β (g)} ≔ \sigma{a ∈ \carr{A(g)}}{\carr{B(g, a)}}\)
  \item \(\eq[Σ A B (g)] {\p{a,b}} {\p{a',b'}} ≔ \eq[A(g)] a {a'} ∧ \eq[B(g,a)] b {b'}\)
  %\item \(\pair(\p{g, a}, b) ≔ \p{g, \p{a,b}} : \ext {\ext Γ A} B ≅ \ext Γ {Σ Α Β}\)
  \item \(\pair(\p{g, a}, b) ≔ \p{g, \p{a,b}}\)
  \end{itemize}
\end{construction}

\begin{construction}
  \(\hvcwf\) has dependent products. Let \(Γ : \ctx\), \(A : \ty Γ\), and \(B : \ty \ext Γ A\).
  \begin{itemize}
  %\item \(Π Α Β (g) ≔
  %  \p{\set{f : \sigma{a ∈ \carr{A(g)}}{\carr{B(g, a)}} → \h}{\ref{hvm:str}-\ref{hvm:unique}},
  %    \p{f, f'} ↦ ⋀_{a ∈ A(g)}⋁_{b ∈ B(g,a)}f(a,b)∧f'(a,b)}\)
  \item \(\carr{Π Α Β (g)} ≔ \set
    {f : \sigma{a ∈ \carr{A(g)}}{\carr{B(g, a)}} → \h}
    {\ref{hvm:str}-\ref{hvm:unique}}\)
  \item \(\eq[Π A B (g)] f {f'} ≔ ⋀_{a ∈ \carr{A(g)}}⋁_{b ∈ \carr{B(g,a)}}f(a,b)∧f'(a,b)\)
  \item \(\tm Γ {\p{Π A B}} = \tm {\p{\ext Γ A}} B\)
  \end{itemize}
\end{construction}


\begin{construction}
  \(\hvcwf\) has unit types \(- ↦ \mb 1 ∈ \fam Γ\).
\end{construction}

\begin{construction}
  \(\hvcwf\) has natural number types \(- ↦ ℕ ∈ \fam Γ\).
\end{construction}

\begin{construction}
  \(\hvcwf\) has extensional identity types.
  \[
    \idtyap[A] a {a'} (g) ≔ \mb 1 \res{\eq[A(g)] a {a'}}
  \]
\end{construction}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End:
